---
title: "A Student's Review of Expert Reviews of Chocolate Bars"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
library(ggplot2)
library(readr)
library(stringr)
library(lubridate)

chocolate <- read.csv("data/flavors_of_cacao[1].csv")
```

# Dataset

The `chocolate` data set contains information about chocolate bars and reviews written by experts. It contains 9 variables with 1795 observations. The data contain information about the company who makes the chocolate bars, the percent of cocoa, where the company is located, bean type & origin, and review (out of 5 stars).  It comes from the kaggle data set, Chocolate Bar Ratings: Extensive EDA, by Will Canniford. The original data can be found [here](https://www.kaggle.com/code/willcanniford/chocolate-bar-ratings-extensive-eda/data.)

We chose this dataset because we believe it had a variety of interesting variables to look at in relationship to the rating the chocolate bar was given. We were especially intrigued by the dataset containing both time and geo-spatial data. We believe that this will allow us to make informative and visually pleasing graphics. We also all love chocolate.

```{r}
#| label: glimpsing-data
head(chocolate, 5)
names(chocolate)

```

# Questions

These are the two questions we want to answer.

\1. What characteristics of the chocolate make up the most successful chocolate bar in reviews?

\2. How do expert reviews change over time for each company?

# Analysis plan

\1. For the first question, we plan to utilize the variables that could effectively distinguish different groups of chocolate bars including the bar brand/name or specific bean origin and graph them against the review ratings to better observe the relationship between the characteristics of chocolates bars and their ratings. In order to do this, we will create a new variables that represent the specific bean origin and the bar name respectively. The "Specific Bean Origin or Bar Name" variable contains varying information and for certain values, we will need to split the character string to obtain data needed.

\2. With regards to the second question, we will make use of the review date and company variables to observe how the expert reviews for different companies vary over time. Then using aesthetic parameters of fill and color with alpha values, we plan to gain a better understanding of what potential changes in characteristics have caused an upward or downward in the review ratings over time.

At this time we do not believe we will need to add any external data to answer either question
